---
title: "Master Thesis - Fetching and processing Data"
author: "Frederick Pena Sims"
format: html
editor: visual
---

## Library

```{r}
library(odbc)
library(RODBC)
library(tidyverse)
library(readxl)
```

## Getting municipal spending data

Explain

```{r}
unique(odbc::odbcListDrivers()[[1]])

dbname <- "C:/Users/frede/Documents/Master Thesis II/master-thesis/Presupuestos2023.accdb"

con <- RODBC::odbcConnectAccess2007(dbname)

RODBC::sqlTables(con)
```

Getting tables I want –\> keeping cdfgr as character to keep preceeding zeroes

```{r}
tb_funcionalPorEntes <- RODBC::sqlFetch(con, "v_funcionalPorEntes",  
                                        as.is = c(F, F, F, F, F, F, F, T, F, F, F))

tb_cuentasProgramas <- RODBC::sqlFetch(con, "tb_cuentasProgramas", as.is = TRUE)

```

## Processing municipal spending data

### Applying official classification

Joining codigo municipio for official classification

Creating common ID

```{r}
codigo_municipio <- read_excel("diccionario24.xlsx", skip = 1, col_names = TRUE) |> 
  mutate(INE = paste0(CPRO, CMUN),
         Provincia = as.integer(CPRO))

tb_funcionalPorEntes <- 
  tb_funcionalPorEntes |>
  mutate(INE = as.character(str_extract_all(CodBDGEL, "^\\d+")))

```

Joining

```{r}
tb_funcionalPorMunicipios <- 
  tb_funcionalPorEntes |> 
  inner_join(codigo_municipio, by = "INE") |> 
  inner_join(tb_cuentasProgramas, by ="cdfgr") |> 
  left_join(tb_provincias, by ="Provincia")
```

### Number of municipalities

```{r}
n_distinct(tb_funcionalPorMunicipios$INE)
```

7426 municipalities

Where is the difference

```{r}
m <-
  tb_funcionalPorMunicipios |>
  mutate(CODAUTO_data = CODAUTO) |> 
  group_by(CODAUTO_data) |> 
  distinct(INE) |> 
  count(CODAUTO_data)

o <- 
  codigo_municipio |>
  mutate(CODAUTO_original = CODAUTO) |> 
  group_by(CODAUTO_original) |> 
  distinct(INE) |> 
  count(CODAUTO_original)

test <- cbind(m, o) |> 
  mutate(difference = n...4 - n...2,
         proportion = (difference/n...4)*100)

test |> 
  ggplot()+
  geom_col(aes(CODAUTO_original, proportion)) +
  geom_text(aes(CODAUTO_original, proportion, label = paste(round(proportion, 2),"%")),
            vjust = -0.4,
            size = 2.5) +
  labs(title = "Share of missing Municipalities by Autonomous Region")

```

### Organising spending data per municipality

Standardizing codification

```{r}

```

Total spending per reason

```{r}
# Group by 'cdfgr' and 'INE', then calculate the (total) sum of spending
summarized_spending <- tb_funcionalPorMunicipios |> 
  group_by(cdfgr, INE) |> 
  summarize(total_spending = sum(importe)) |> 
  ungroup()

total_importePorMunicipio <- tb_funcionalPorMunicipios |> 
  group_by(INE) |> 
  summarize(total_importe = sum(importe)) |> 
  ungroup()
```

Keeping only one spending amount and spending reason per municipality

```{r}
tb_importePorPrograma <-
  tb_funcionalPorMunicipios |> 
  left_join(summarized_spending, by = c("cdfgr", "INE")) |>
  left_join(total_importePorMunicipio, by = "INE") |> 
  group_by(INE, cdfgr) |> 
  distinct(cdfgr, .keep_all = T) |> 
  ungroup()
```
